---
title: "Homework 03"
author: "Emilio Ojemann"
date: 2026-05-27
execute:
  warning: false
  message: false
  fig-width: 4
  fig-height: 5
format: 
  pdf:
    toc: true
---

# Part 1. Set up tasks

## GitHub steps
- Created repository on GitHub
- Added description, made public, contains a README

## RStudio steps

**Read in packages**

```{r}

library(tidyverse)
library(here)
library(janitor)
library(readxl)

```


**Store kelp data**

```{r}

kelp <- read_csv("../data/temp-kelp.csv")

```

**Store personal data**




# Part 2. Problems

## Problem 1. Giant kelp fronds

**a. An appropriate test**

To determine the strength of the relationship between temperature and giant kelp frond elongation rate, I would use either a Pearson correlation or a 

**b. Create a visualization**

**c. Check your assumptions and run your test**

**d. Results communication**

**e. Test implications*

**f. Double check your own work**


## Problem 2. Personal data

## Problem 3. Affective visualization

## Problem 4. Statistical critique

