---
title: "Homework 03"
author: "Emilio Ojemann"
date: 2026-05-27
execute:
  warning: false
  message: false
format: 
  pdf:
    toc: true
---

# Part 1. Set up tasks

## GitHub steps
- Created repository on GitHub
- Added description, made public, contains a README

## RStudio steps

**Read in packages**

```{r}

library(tidyverse)
library(here)
library(janitor)
library(readxl)

```


**Store kelp data**

```{r}

kelp <- read_csv("../data/temp-kelp.csv")

```

**Store personal data**




# Part 2. Problems

## Problem 1. Giant kelp fronds

**a. An appropriate test**

To determine the strength of the relationship between the ocean temperature and giant kelp frond elongation rate, two viable test would be either a Pearson correlation or Spearman's rank correlation.

Pearson's correlation helps to describe the strength and direction of an assumed linear relationship between two continuous, normally distributed variables. It additionally assumes independent observations. The correlation coefficient "r" is specifically helpful to describe how much of the variance in elongation rate is attributed to temperature.

Spearman's rank correlation is the non-parametric version, making no assumption about the distribution of the variables. The data is treated as ranked, so only a monotonic relationship is needed and, just like the Pearson's correlation, requires independent observations. 


**b. Create a visualization**

```{r}

# base layer: ggplot
ggplot(data = kelp, # kelp data frame
       aes(x = temp_c, # x-axis: temperature
           y = kelp_elong)) + # y-axis: elongation rate
  geom_point(color = "green4", # non-default color
             size = 1.5, # size of points
             alpha = 0.7) + # transparency of points
  # title and axis labels
  labs(title = "As temperature increases, kelp frond decreases", # informative title
       x = "Temperature (˚C)", # x axis label with units
       y = "Kelp elongation (cm/day)") + # y axis label with units
  theme_classic() # non-default theme


```

**c. Check your assumptions and run your test**

```{r}

# fitting a linear model
kelp_model <- lm(kelp_elong ~ temp_c, # response ~ predictor
                 data = kelp) # data frame 

# checking assumptions
par(mfrow = c(2,2)) # 2x2 grid
plot(kelp_model) # residuals v fitted, Q-Q, scale-location, leverage
par(mfrow = c(1,1)) # reset grid


```
I checked the plots above for several assumptions after fitting the model. First, I assessed linearity of points in the 1b visualization; there is a visible negative relationship between temperature and elongation rate. 

Homoscedasticity is checked in the with the Residuals v. Fitted and Scale-Location plots. The residuals are evenly scattered around the dotted line in the former and latter. 

Then, normality of residuals is assessed with Q-Q plot. The points in the plot generally follow the diagonal reference line. There is some deviation toward the ends of the line, though the normality assumption is generally met. 


```{r}

cor.test(kelp$temp_c,
         kelp$kelp_elong,
         method = "pearson")

```

**d. Results communication**

To evaluate the strength of the relationship between temperature and giant kelp frond elongation rate, I used a Pearson's correlation, because both variables are continuous and the initial visualization demonstrated a linear relationships. Other assumptions like normally distributed residuals and variance were met - and checked after fitting the model.

There was a significant negative relationship between ocean temperature and giant kelp frond elongation rate (Pearson's r= -0.69, t(30)= -5.17, p<0.001, $\alpha$=0.05), indicating a moderate negative association: as ocean temperature increased, kelp frond elongation rate decreased. 


**e. Test implications**

The analysis indicates a significant negative relationship between ocean temperature and giant kelp frond elongation rate (Pearson's r = -0.69, p <0.001): when ocean temperature increases, kelp fronds tend to grow more slowly. In the broader context of the climate crisis, warmer oceans could threaten the health and general density of kelp forests as their growth rates decline. Down the line, any and all species dependent on these forests for food or habitat will be impacted.

**f. Double check your own work**




## Problem 2. Personal data

## Problem 3. Affective visualization

## Problem 4. Statistical critique

